Mainstreaming Race Science

### 2023-10-27 paper/paper.qmd
small conversion errors

format: 
  elsevier-pdf:
    journal: 
      name: "journal"
    keep_tex: true
    latex_engine: "xelatex"

@@

# Introduction {.unnumbered}

@@

# Scientific racism, eugenics, and *Brown* {.unnumbered}

@@

# *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}

@@

# Topic model analysis identifies race science discourse {.unnumbered}

@@

# Race science, behavior genetics, and psychology {.unnumbered}

@@

\clearpage
\pagenumbering{arabic}
\renewcommand*{\thepage}{\thesection.\arabic{page}}
\appendix
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand\thefigure{\thesection.\arabic{figure}}    
\setcounter{figure}{0} 
\renewcommand\thetable{\thesection.\arabic{table}}    
\setcounter{table}{0}

@@

@@

![Corpus assembly.](img/dataset.png){#fig-corpus width="8in" height="6.1in"}

@@

An electronic archive of *Mankind Quarterly*, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.  The Python package Beautiful Soup (<https://www.crummy.com/software/BeautifulSoup/>, version not recorded) was used to retrieve every PDF available in this archive.  After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.  Text was extracted using the R package `pdftools` \cite[version 3.0.1]{OomsPdftoolsTextExtraction2023}.  This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.

@@

To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.  Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda [@TuckerFundingScientificRacism2002].  After the *Brown v. Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005}.  We reviewed a critical profile of Pioneer [MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.

@@

The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.  We used the Crossref API and `rcrossref` R interface to this API \cite[version 1.1.0.99]{ChamberlainRcrossrefClientVarious2019] to retrieve metadata for each item published from 1960-2010 in each of these journals.  These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.  These URLs were used to retrieve a XML or PDF version of each item, except for *Personality and Individual Differences*.  This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.  (This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.  One set of examples is an exchange between Rushton and Flynn on the Flynn effect \cite{FlynnEvidenceRushtonGenetic1998,RushtonSecularGainsIQ1998,FlynnReplyRushtonGang1998}.)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.  This stage of corpus assembly was conducted between 2021-11 and 2022-05.

@@

After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.  @fig-counts and table @tbl-counts show the number of fulltext documents included in the corpus.  All together, the corpus includes 34,896 documents from 6 journals.

@@

After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library \cite[version 1.9.0]{SpaCyIndustrialstrengthNatural2018} and the R API `spacyr` \cite[version 1.2.1]{BenoitSpacyrWrapperSpaCy2020}.  Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").  We then counted the occurrence of each noun phrase in the document.  The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R \cite[version 7.0.0]{RichardsonArrowIntegrationApache2023}.

@@

At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019} and human geneticists \cite{JacksonFacingOurHistory2023].  Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].  A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.  A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," but primarily focuses on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).

>>>>

format: 
  elsevier-pdf:
    journal: 
      name: "journal"
    keep_tex: true
    latex_engine: "xelatex"
    number-sections: false

@@

<!-- # Introduction {.unnumbered} -->

@@

<!-- # Scientific racism, eugenics, and *Brown* {.unnumbered} -->

@@

# Results

## *Mankind Quarterly* and Pioneer-funded researchers {.unnumbered}

@@

## Topic model analysis identifies race science discourse {.unnumbered}

@@

## Race science, behavior genetics, and psychology {.unnumbered}

@@

<!-- 
\clearpage
\pagenumbering{arabic}
\renewcommand*{\thepage}{\thesection.\arabic{page}}
\appendix
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand\thefigure{\thesection.\arabic{figure}}    
\setcounter{figure}{0} 
\renewcommand\thetable{\thesection.\arabic{table}}    
\setcounter{table}{0}

@@

\clearpage
 -->

@@

![Corpus assembly.](img/dataset.png){#fig-corpus width="120%"}

@@

An electronic archive of *Mankind Quarterly*, from the first issue in 1969 through 2004, is available for free on the open web at the white nationalist website *The Unz Review*.  The Python package Beautiful Soup (<https://www.crummy.com/software/BeautifulSoup/>, version not recorded) was used to retrieve every PDF available in this archive.  After identifying some gaps (missing issues) in the archive from *The Unz Review*, we manually retrieved PDFs for further articles using ProQuest.  Text was extracted using the R package `pdftools` [@OomsPdftoolsTextExtraction2023 version 3.0.1].  This stage of corpus assembly was conducted in Fall 2021 with the assistance of Anthony Sainez.

@@

To identify suitable mainstream journals for inclusion in the corpus, we first identified academic researchers who had been funded by the Pioneer Fund.  Pioneer is an American non-profit organization founded in 1937 to fund eugenics research and propaganda [@TuckerFundingScientificRacism2002].  After the *Brown v. Board of Education* ruling in 1954, Pioneer funded various segregationist efforts across the United States, including lectures on eugenics by Stanford physicist William Shockley [@JacksonScienceSegregationRace2005].  We reviewed a critical profile of Pioneer [@MillerPioneerFundBankrolling1994] as well as an archived page from the organization's own web site (<https://web.archive.org/web/20130103005545/http://www.pioneerfund.org/Grantees.html>), which together listed 16 researchers who had received Pioneer funds.

@@

The remaining 5 journals are all published by major academic publishers — Elsevier, Springer, or Cambridge University Press — and each item in the entire run of each journal has been assigned a DOI (digital object identifier) for archival purposes.  We used the Crossref API and `rcrossref` R interface to this API [@ChamberlainRcrossrefClientVarious2019 version 1.1.0.99] to retrieve metadata for each item published from 1960-2010 in each of these journals.  These metadata included item-level license information — confirming that the text of each item could be used for text mining projects — and a URL to an electronic version of the item.  These URLs were used to retrieve a XML or PDF version of each item, except for *Personality and Individual Differences*.  This journal has published a relatively large number of non-article documents, such as book reviews and commentaries, that are not available at the URL included in the Crossref metadata.  (This is unfortunate, as it was not difficult to find highly relevant documents that we could not automatically retrieve and therefore had to be excluded from our corpus.  One set of examples is an exchange between Rushton and Flynn on the Flynn effect [@FlynnEvidenceRushtonGenetic1998; @RushtonSecularGainsIQ1998; @FlynnReplyRushtonGang1998].)  Instead we used Elsevier's ScienceDirect API (<https://dev.elsevier.com/>) to independently search and retrieve all available items from *Personality and Individual Differences*.  This stage of corpus assembly was conducted between 2021-11 and 2022-05.

@@

After XML/PDF versions were retrieved, text was extracted using either the `xml2` package [@WickhamXml2ParseXML2023] or `pdftools`.  @fig-counts and @tbl-counts show the number of fulltext documents included in the corpus.  All together, the corpus includes 34,896 documents from 6 journals.

@@

After document retrieval and text extraction, we pre-processed the text using the `spaCy` NLP (natural language processing) Python library [@SpaCyIndustrialstrengthNatural2018 version 1.9.0] and the R API `spacyr` [@BenoitSpacyrWrapperSpaCy2020 version 1.2.1].  Specifically, we applied regular expressions to remove header/footer copyright notices and hyphenation, used spaCy to annotate and extract noun phrases (eg, "the intelligence test items"), and then cleaned and standardized these phrases (eg, removing the/an/a, coverting all text to lowercase, and replacing all whitespace with underscores: "intelligence_test_items").  We then counted the occurrence of each noun phrase in the document.  The aggregated "document-term matrix" was stored in Parquet format [@VohraApacheParquet2016] for performance reasons, and written and read using the `arrow` package for R [@RichardsonArrowIntegrationApache2023 version 7.0.0].

@@

At the same time, we were unable to find any articles published in BG that acknowledged the involvement of the field with scientific racism with anything like the force of the statements that have been made by biological anthropologists [@FuentesAAPAStatementRace2019] and human geneticists [@JacksonFacingOurHistory2023].  Notably, the ASHG report includes discussions of behavior genetics race science, and in particular is critical of the field and the organization for failing to publicly reject the race-and-intelligence claims made by Shockley and Jensen [@JacksonFacingOurHistory2023].  A search for "eugenics" in BG found articles that presented the issue as part of the distant past, and a search for "racism" turned up studies of racist attitudes.  A statement on the Behavior Genetics Association website, dated 2021, acknowledges that "The history of our field is inextricably linked with racism, including the misuse of behavior genetic research to support violent eugenic policies," but primarily focuses on BGA member demographics (<https://www.bga.org/content.aspx?page_id=22&club_id=971921&module_id=567723>).
